---
title: "21 July 2022 - 29 July 2022"
author: "Eric O'Malley"
date: "29 July 2022"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{The neotoma2 R Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
```

```{r setup, message=FALSE}
library(neotoma2)
library(dplyr)
library(purrr)
library(sf)
library(tidyverse)
library(rioja)
library(leaflet)
library(leaflegend)
```

Work done over 21 July 2022 - 29 July 2022

# OVERVIEW / ACCOMPLISHED


- 1. Successful (I think) stratigraphic plotting (multiple taxa one site AND one taxa multiple sites). Solid understanding of manipulating the graph using **ggplot**. Stratigraphic plot made using **rioja** as well, unsure how to interpret. The rioja ones still have some issue with formatting the visuals.
  - a. See Question 1. Easy fix to incorporate
  
- 2. While going through the EPD Binder examples, there was a section on taxa harmonization. I've modified that code for a diatom example and placed it within the "Diatom_Taxa.Rmd"

- 3. Also from the EPD Binder, useful code on transforming data (obtained from `samples()`) from the long format its given to a wide format. The code is used in the data retrieval steps for the stratigraphic diagrams, but I've also included this code within the code I started just prior to your vacation, which was data retrieval using an aggregate dataset. The code now starts from an excel file **to** having count data in both long and wide formats.

- 4. Leaflet -- Made an example map plotting two different variables (abundance & pH). I'm unsure of how to make it visually appealing with enough of a contrast between the two. I've figured out some things that I think can help me with the SHINY app.

- 5. Added some code chunks to Diatoms_analysis.Rmd & Diatoms_neo2.Rmd that I thought were useful. Mostly involved data transformation. Did not add any of the below code into those R markdowns yet.



# QUESTIONS

- 1. Percent Abundance: This should be calculated based each sample? I just need clarification for how I should approach this for the stratigraphic diagrams. Should I calculate percent abundance as soon as I get the output from `samples()`? So essentially all the counts in one sample (a single sampleid) should add up to 100%? I know we had this down, but i've been looking at too many pollen examples and I'm all messed up.

- 2. As you can see in the ggplot strat example, I chose 4 sites that all contained 1 taxa I manually found in the 4 datasets. The counts varied wildly which is not good for comparing in these stratigraphic diagrams. Did Simon say what the best way to solve this was?

# EPD BINDER

## Taxa Harmonization

I've made changes to the "Diatom_taxa.Rmd" to include code on Taxa Harmonization. This code was obtained from the EPD Binder. I know this was not on the list, but it was one of the first steps in the process Simon goes through.

### Data Retrieval

The following code chunk is used in most of the code below

```{r data retrieval, warning=FALSE}
# polygon obtained from http://geojson.io/#map=5/42.049/-75.894
ne <- '{"type": "Polygon",
        "coordinates": [[
            [-80.5517578125,
              39.70718665682654],
            [-66.7529296875,
              39.70718665682654],
            [-66.7529296875,
              47.517200697839414],
            [-80.5517578125,
              47.517200697839414],
            [-80.5517578125,
              39.70718665682654]]]}'


ne_datasets <- neotoma2::get_datasets(loc = ne[1], datasettype = "diatom",all_data = TRUE)

ne_dl <- get_downloads(ne_datasets, all_data = TRUE)

allSamp <- samples(ne_dl)

```


### Taxa Harmonization Step

Code modified from Simon Goring and Socorro Dominguez Vidaña's [EPD Binder: Simple Workflow](https://open.neotomadb.org/EPD_binder/simple_workflow.html#)


```{r, taxa harmonization, warning=FALSE}
allSamp <- allSamp %>%
  mutate(variablename = replace(variablename, 
                                stringr::str_detect(variablename, "lirata.*"), 
                                "Aulacoseira lirata"))
#`mutate()` the column `variablename` so that any time we detect (str_detect()) a variablename that has lirata" (the .* represents a wildcard for any character [.], zero or more times [*]) we replace() it with the character string "Aulacoseira lirata"

original_taxa_list <- neotoma2::taxa(ne_dl)
# `taxa()` can be used to get the original list of taxa to compare changes.

changed <- unique(allSamp$variablename)
original <- original_taxa_list$variablename

setdiff(original, changed)
#displays what taxa were changed
```

##Data Manipulation

```{r, data manipulation for diagrams}
#Get a singular site, select only taxa identified from diatom cores
## In this case its Upper Wallface Pond
plottingSite <- ne_dl[[1]]

# "plottingTaxa" is the taxa we want to plot. The code currently gets the 10 taxa that appears in the most samples
# (REPLACE WITH FILTERING BY PERCENT ABUNDANCE (IE: WHEN ABUNDANCE IS >5%)) CODE FOUND IN OLDER STRATIGRAPHIC CODE
plottingTaxa <- neotoma2::taxa(plottingSite) %>%
  arrange(desc(samples)) %>% 
  head(n = 10)

# The counts from Upper Wallface Pond are filtered to only have information from
# the taxa we determined that we wanted in "plottingTaxa"
shortSamples <- samples(plottingSite) %>% 
  filter(variablename %in% plottingTaxa$variablename)

# Transform to proportion values. 
# **IM UNSURE IF WE SHOULD DO THIS FOR DIATOMS (PERCENT ABUNDANCE INSTEAD?)**
onesite <- shortSamples %>%
  group_by(age) %>%
  mutate(diatomcount = sum(value, na.rm = TRUE)) %>%
  group_by(variablename) %>% 
  mutate(prop = value / diatomcount) %>% 
  arrange(desc(age))

# Spread the data to a "wide" table, with taxa as column headings.
widetable <- onesite %>%
  dplyr::select(age, variablename, prop) %>%
  mutate(prop = as.numeric(prop))

counts <- tidyr::pivot_wider(widetable,
                             id_cols = age,
                             names_from = variablename,
                             values_from = prop,
                             values_fill = 0)
```

## Rioja Stratigraphic Plot

The following `strat.plot()` example is based off of the EPD binder

```{r, rioja strat, warning=FALSE, out.width = '100%'}
clust <- rioja::chclust(dist(sqrt(counts)),
                        method = "coniss")
plot <- rioja::strat.plot(counts[,-1] * 100, yvar = counts$age,
                  title = ne_dl[[1]]$sitename,
                  ylabel = "Calibrated Years BP",
                  xlabel = "Diatom (%)",
                  y.rev = TRUE,
                  clust = clust,
                  wa.order = "topleft", scale.percent = TRUE)

```

The following `strat.plot()` example is based off of Benjamin Bell's Blog

```{r, another rioja example, out.width = '100%'}
#https://www.benjaminbell.co.uk/2018/02/pollen-diagrams-in-r-using-rioja-part-1.html
#code chunk same as above EXCEPT for not adding "prop" variable and leaving it as NISP counts

onesite2 <- shortSamples %>%
  arrange(desc(age))

widetable2 <- onesite2 %>%
  dplyr::select(age, variablename, value)

counts2 <- tidyr::pivot_wider(widetable2,
                             id_cols = age,
                             names_from = variablename,
                             values_from = value,
                             values_fill = 0)

#below is the Rioja plot that I could get to produce the best plot after fiddling with it for some time.

strat.plot(counts2[,-1], yvar = counts$age, y.rev=TRUE, plot.line=FALSE, plot.poly=TRUE, plot.bar=TRUE, col.bar="black", col.poly="green", col.poly.line="black", scale.percent=TRUE, xSpace=0.01, x.pc.lab=TRUE, x.pc.omit0=TRUE, las=2,srt.xlabel=25,title="Upper Wallface Pond Diatoms", ylabel="Age (cal yr BP)",cex.yaxis=0.6, cex.axis=0.6,cex.xlabel=0.8,ylabPos=3, wa.order = "topleft")

#ISSUES: col.poly.line="black" does not work. I assume no bottom x label because they are expected to be in percents? Awkward angle for taxa labels, only way to avoid hitting title.



#To save to view the whole graph: Go to Export (in plot viewer) -> Save as PDF -> 14 in by 14 in (these dimensions seem to work best after messing around with it)
```

## GGPLOT - Multiple Sites (1 taxa) 

By using what I've learned from above using mostly Simon's code (altered for Diatoms) I will try and create the plot for multiple sites.

```{r, multiple sites strat, out.width = '100%'}
#plottingSites <- ne_dl[[1:5]]

plottingSite1 <- ne_dl[[1]]
plottingSite2 <- ne_dl[[2]]
#Site 3 didn't seem like a good example
plottingSite4 <- ne_dl[[4]]
plottingSite5 <- ne_dl[[5]]

#view(neotoma2::taxa(plottingSite1))
#view(neotoma2::taxa(plottingSite2))
#view(neotoma2::taxa(plottingSite4))
#view(neotoma2::taxa(plottingSite5))

#After viewing the taxa content of the 4 sites, I found that Achnanthes marginulata
# was present in all 4

# Now I will filter each site by Achnanthes marginulata

UWall_AM_counts <- samples(plottingSite1) %>% 
  dplyr::filter(variablename == "Achnanthes marginulata")

Deep_AM_counts <- samples(plottingSite2) %>%
  dplyr::filter(variablename == "Achnanthes marginulata")

Georgia_AM_counts <- samples(plottingSite4) %>%
  dplyr::filter(variablename == "Achnanthes marginulata")

Blue_AM_counts <- samples(plottingSite5) %>%
  dplyr::filter(variablename == "Achnanthes marginulata")

ALL_AM_counts <- rbind(UWall_AM_counts, Deep_AM_counts, Georgia_AM_counts, Blue_AM_counts)

```

```{r}
ggplot_mult <- ALL_AM_counts %>%
  ggplot(aes(y = round(ALL_AM_counts$age, digits = 0), x = value)) +
  geom_path() + #adds lines between each value
  geom_point() + #adds points
  facet_grid(~sitename, scales = "free_x", space = "free_x") + #`facet_wrap()` is what each graph will represent. `scales` is used to give each graph a its own scaled axis (default will scale every graph to same x axis). "free_x" and "free_y" are used seperately to seperate the x axis and y axis respectively.
  scale_y_reverse() +
  labs(x = "Count (NISP)", y = "Age (cal yr BP)") +
  theme_bw() +
  theme(
    # rotate the facet labels
    strip.text.x = element_text(angle = 45, hjust = 0, vjust = 0), 
    # turn off the label background
    strip.background = element_blank(),
    axis.line = element_line(color='black'),
    #plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    #panel.border = element_blank()
  )

ggplot_mult
```

```{r}
###########################################################
species_plot_grob <- ggplotGrob(ggplot_mult)
for(i in which(grepl("strip-t", species_plot_grob$layout$name))){
  species_plot_grob$grobs[[i]]$layout$clip <- "off"
}

# needed to draw the modified plot_grob
grid::grid.draw(species_plot_grob)
```

 I use this function of mine in the next chunks
```{r percent abundance function, echo=FALSE}
pc_abun <- function(dl_object) {
  
  samp <- samples(dl_object)
  taxa <- samp$variablename
  count <- samp$value
  sitename <- samp$sitename
  datasetid <- samp$datasetid
  siteid <- samp$siteid
  sampleid <- samp$sampleid
  lat <- samp$lat
  long <- samp$long
  
  
  countDF <- data.frame(lat, long, siteid, sitename, datasetid, sampleid, taxa, count) %>%
    group_by(sampleid) %>%
    mutate(count_pct = count*100 / sum(count))
  return(countDF)
}
```



# leaflet

The goal of the code below is to not only show percent abundance and location, but add other variables (like WC variables). Previously I tried varying the size by abundance and the color by WC characteristic but that didn't work. Here I will try adding either:
- 2 different markers that would not clash (visually confusing)
- A different approach using conditional statements. If the characteristic is above     a certain level (maybe something deemed high) the color will change. This works     differently than my previous attempt due to continuous vs categorical values
    being treated differently.

I also want to add more options to the popup box, such as actual abundance and WC values. This could also be used for the SHINY app. 


```{r, two variable leaflet}
#This is using the "EVERYTHING" sheet in the aggregare dataset as of 7-20-22
DREXEL_agg <- read.csv("Aggregate Dataset CSV 7-29-22.csv")
colnames(DREXEL_agg) <- c("drexelid","siteid","sitename","diatom_datasetid", "wc_datasetid")

with_wc_agg <- DREXEL_agg[!is.na(DREXEL_agg$wc_datasetid),]

diatom_datasetid <- tail(with_wc_agg$diatom_datasetid,50) #for some reason I got an error trying to do `head()` instead of `tail()`. Possible wrong datasetid towards beginning.

wc_datasetid <- tail(with_wc_agg$wc_datasetid,50)

diatom_dl <- get_datasets(diatom_datasetid,all_data = TRUE) %>%
  get_downloads()

wc_dl <- get_datasets(wc_datasetid,all_data = TRUE) %>%
  get_downloads()

diatom_taxa <- pc_abun(diatom_dl) %>%
  dplyr::filter(taxa == "Achnanthes minutissima") %>%
  arrange(siteid)

wc_variable <- samples(wc_dl) %>%
  select('siteid', 'variablename','value') %>%
  dplyr::filter(variablename == "pH") %>%
  arrange(siteid)

wc_diatom <- merge(diatom_taxa, wc_variable)
```


```{r, editing plotleaflet function, out.width = '100%'}
getColor <- function(wc_diatom) {
  sapply(wc_diatom$value, function(value) {
  if(value <= 8.2) {
    "red"
  } else {
    "blue"
  }})
}

symbols <- makeSizeIcons(values = wc_diatom$count_pct,
                          shape = 'circle',
                          color = 'black',
                          fillColor = getColor(wc_diatom),
                          #pal = numPal,
                          opacity = .7,
                          fillOpacity = .5,
                          baseSize = 15)
            
             map1 <- leaflet(wc_diatom) %>%
              addProviderTiles(providers$Stamen.TerrainBackground) %>%
              addTiles() %>%
              addMarkers(icon = symbols, lng = wc_diatom$long, lat = wc_diatom$lat,
                               popup = paste0("<b>", wc_diatom$sitename,
                                              "</b><br><b>pH:</b> ",
                                              wc_diatom$value,
                                              "<br><a href=http://apps.neotomadb.org/explorer/?siteids=",
                                              wc_diatom$siteid,
                                              ">Explorer Link</a>"),
                               options = markerOptions(riseOnHover = TRUE)) %>%
              addLegendSize(values = wc_diatom$count_pct,
                fillColor = c('white'),
                color = 'black',
                #pal = numPal,
                title = "A. minutissima % Abundance",
                baseSize = 15,
                shape = 'circle',
                orientation = 'horizontal',
                opacity = .7,
                fillOpacity = .3,
                position = 'bottomright',
                breaks = 5) #%>%
               # addLegend(values = c())
             
map1
```





```{r, adding icon for WC variable, out.width = '100%'}
numPal <- colorNumeric('magma', wc_diatom$value)

symbols2 <- makeSizeIcons(values = wc_diatom$value,
                                     shape = 'plus',
                                     color = "black",
                                     pal = numPal,
                                    #fillColor = ifelse(wc_diatom$value <= 8.0,'blue','red'),
                                     opacity = 1,
                                     baseSize = 12)
  map1 %>%
    addMarkers(icon = symbols2, lng = wc_diatom$long, lat = wc_diatom$lat, popup = paste0("<b>", wc_diatom$sitename,
                                              "</b><br><b>pH:</b> ",
                                              wc_diatom$value,
                                              "<br><a href=http://apps.neotomadb.org/explorer/?siteids=",
                                              wc_diatom$siteid,
                                              ">Explorer Link</a>"),
                               options = markerOptions(riseOnHover = TRUE)) %>%
    #Cannot get a working legend for WC
    
    addLegendSize(values = wc_diatom$value,
                #color = "black",
                pal = numPal,
                #fillColor = ifelse(wc_diatom$value <= 8.0,'blue','red'),
                title = "pH",
                shape = 'plus',
                orientation = 'horizontal',
                opacity = 1,
                fillOpacity = 1,
                position = 'bottomright',
                breaks = 2,
                baseSize = 12)
```


#GGPLOT FACETED STRATIGRAPHIC PLOTS

This code also contains all of the filtering steps we discussed.

```{r, more in depth stratigraphic diagram filtering}
# STEP 1: Getting data out of NEOTOMA

test <- get_datasets(datasettype = "diatom", all_data = TRUE)

core_1 <- get_datasets(datasettype = "diatom", limit = 300) %>%
  get_downloads()

core_1_samp <- samples(core_1)

core_2 <- get_datasets(datasettype = "diatom", limit = 300, offset = 301) %>%
  get_downloads()

core_2_samp <- samples(core_2)

core_3 <- get_datasets(datasettype = "diatom", limit = 300, offset = 602) %>%
  get_downloads()

core_3_samp <- samples(core_3)

core_4 <- get_datasets(datasettype = "diatom", limit = 179, offset = 903) %>%
  get_downloads()

core_4_samp <- samples(core_4)

#siteid: 26515
#will not download, skip

core_5 <- get_datasets(datasettype = "diatom", limit = 300, offset = 1100) %>%
  get_downloads()

core_5_samp <- samples(core_5)

all_diatoms <- rbind(core_1_samp,core_2_samp,core_3_samp,core_4_samp,core_5_samp)

#twenty_diatoms <- get_datasets(datasettype = "diatom", limit = 20) %>%
  #get_downloads() %>%
  #samples() #BECAUSE ALL_DIATOMS TOO LARGE, TESTING THIS.
#IF TO GO BACK, REPLACE twenty_diatoms with all_diatoms
###############################################################################
# OPTIONAL STEP: COMBINING WATER CHEMISTRY DATA
core_sites <- unique(all_diatoms$siteid) %>%
  get_sites()

core_wc <- get_datasets(core_sites)
# %>% neotoma2::filter(datasettype == ""water chemistry")

wc_samp <- get_downloads(core_wc) %>%
  samples() %>%
  dplyr::filter(datasettype == "water chemistry") %>%
  arrange(.,siteid) %>%
  select(-datasetnotes,-database,-taxongroup,-ecologicalgroup,-symmetry,-context,-sitenotes,-description)

#colnames(wc_samp) <- c("siteid","water_chemistry","wc_units","wc_value")

some_diatoms_less <- some_diatoms %>%
  arrange(.,siteid) %>%
  select(-datasetnotes,-database,-taxongroup,-ecologicalgroup,-symmetry,-context,-sitenotes,-description)

wc_diat <- rbind(wc_samp, some_diatoms_less) %>%
  arrange(.,siteid)

testing <- wc_samp %>%
  group_by(siteid) %>%
  summarize(length(unique(datasetid)))

test2 <- list(wc_samp,some_diatoms_less)

#test <- merge(wc_samp,some_diatoms_less, by = "siteid")



###############################################################################
#STEP 2: Filtering out TOP - BOTTOM samples
all_diatom_no_samp <- all_diatoms %>%
  group_by(datasetid) %>%
  summarize(no_samples = length(unique(sampleid)))

ind <- (all_diatom_no_samp$no_samples <= 3)
okay <- all_diatom_no_samp[!ind, ]

newdf <- all_diatoms[all_diatoms$datasetid %in% okay$datasetid,]

###############################################################################
#STEP 3: Percent Abundance calculation
newdf2 <- newdf %>%
  group_by(sampleid) %>%
  mutate(count_pct = value*100 / sum(value))

newdf3 <- newdf2[complete.cases(newdf2$age),] #gets rid of any rows that contain NAs in age column

###############################################################################
# Step 4: Taxa Harmonization
newdf4 <- newdf3 %>%
  mutate(variablename = replace(variablename, 
                                stringr::str_detect(variablename, "stelligera.*"), 
                                "Discostella stelligera"))

original_taxa <- unique(newdf3$variablename)
new_taxa <- unique(newdf4$variablename)

setdiff(original_taxa, new_taxa)
#[1] "Cyclotella stelligera"           "Cyclotella pseudostelligera"    
#[3] "Discostella pseudostelligera"    "Discostella stelligera var. ?"  
#[5] "Cyclotella stelligera (complex)"
###############################################################################
# Step 5: Combining % Abundance values after taxa harmonized

newdf5 <- newdf4 %>%
  dplyr::filter(variablename == "Discostella stelligera")

#n_occur <- data.frame(table(newdf5$variablename))
#n_occur[n_occur$Freq > 1,] #useful if filtering for more than one taxa. Takes only taxa that occur in more than one samples

ok <- aggregate(value ~ sampleid + variablename, FUN = sum, data = newdf5)

test <- merge(ok,newdf5,by = "sampleid")

test2 <- test %>%
  select(-value.y, -variablename.y)

#test3 <- test2[!duplicated(test2),]

test3 <- test2[!duplicated(test2[c('sampleid')]), ] %>%
  rename(value = value.x) %>%
  rename(variablename = variablename.x)

##############################################################################
# Step 6: Filters out "collection date" age models
testing <- test3 %>%
  group_by(datasetid) %>%
  distinct(age)

testing2 <- testing %>%
  filter(length(datasetid) == 1)

testing2 <- as.numeric(testing2$datasetid)

new_test3 <- test3 %>%
  dplyr::filter(!(datasetid %in% testing2))

###############################################################################
# Step #: changing to calendar years instead of BP

new_test4 <- new_test3 %>%
  mutate(age = 1950-age)
  
###############################################################################
# Step 7: filtering for a fixed range of ages

fixed_1 <- new_test4 %>%
  dplyr::filter(new_test3$age >= 1700 & new_test3$age <= 2020)

#fixed_2 <- head(fixed_1, 200)

fixed_2 <- fixed_1 %>%
  dplyr::filter(!(datasetid == 22305))
  

eyee <- fixed_2 %>%
  group_by(sitename) %>%
  summarise(length(unique(datasetid))) #checking to see what sites have more than 1 diatom core dataset

##############################################################################
# Step #: Modifying data so that Strat shows 0% abun at other ages counted

newdf4$NEW <- ifelse(newdf4$variablename != "Discostella stelligera", 0, newdf4$count_pct)

okayy <- newdf4 %>%
  group_by(datasetid) %>%
  summarize(unique(age))

#colnames(okayy) <- c("datasetid", "age")

okayy2 <- merge(okayy, fixed_2, all.x = TRUE)

okayy3 <- okayy2 %>%
  group_by(datasetid) %>%
  mutate(count_pct = ifelse(is.na(count_pct), 0, count_pct))

  
okayy4 <- okayy3 %>%
  group_by(datasetid) %>%
  summarise(sum(count_pct))

colnames(okayy4) <- c("datasetid", "count_pct")

okayy5 <- okayy4 %>%
  dplyr::filter(count_pct == 0)

okayy_fin <- okayy3 %>%
  dplyr::filter(!(datasetid %in% okayy5$datasetid)) 

okayy_fin_2 <- okayy_fin %>%
  dplyr::filter(age >= 1700 & age <= 2020)


#########################################################################
# Step #: Filtering for sites using a max abundance threshold
site_max_abun <- okayy_fin_2 %>%
  group_by(sitename) %>%
  summarise(max_abund = max(count_pct)) %>%
  filter(max_abund >= 7.3) %>%
  arrange(max_abund) %>%
  pull(sitename)

okayy_fin_max <- okayy_fin_2 %>%
  filter(sitename %in% site_max_abun) %>%
  mutate(sitename = factor(sitename, levels = site_max_abun)) %>%
  arrange(sitename)

##############################################################################
# THIS MAKES THE GGPLOT
ggplot_stelligera <- okayy_fin_2 %>%
  ggplot(aes(y = round(age, digits = 0), x = count_pct)) +
  ggtitle("Discostella stelligera") + #adds title
  geom_path() + #adds lines between each value
  geom_point() + #adds points
  facet_wrap(~sitename, scales = "free_x") + #`facet_wrap()` is what each graph will represent. `scales` is used to give each graph a its own scaled axis (default will scale every graph to same x axis). "free_x" and "free_y" are used seperately to seperate the x axis and y axis respectively.
  scale_y_reverse() +
  labs(x = "Abundance (%)", y =
         #"Age (cal yr BP)"
         "Age") +
  theme_bw()


ggsave('stelligera9.jpeg')

# ggsave('stelligera8.jpeg', height = 100, width = 13, units = "in", limitsize = FALSE) # for a large amounts of facets where u can scroll the image



##############################################################################
#setdiff(test$value.x,test$value.y)

#newdf5 <- newdf4 %>%
  #group_by(sampleid) %>%
  #aggregate(count_pct ~ variablename, sum)

#newdf5 <- aggregate(newdf4$count_pct, by = list(newdf4$sampleid, newdf4$variablename), FUN = sum)

#newdf6 <- merge(newdf5, newdf4, all.y = TRUE)
  #aggregate(count_pct, by = list(as.numeric(variablename)), FUN = sum, drop = FALSE)

#df4_filt <- newdf4 %>%
  #dplyr::filter(variablename == "Discostella stelligera")

#df5_filt <- newdf5 %>%
  #dplyr::filter(variablename == "Discostella stelligera")

#filter out datasets with 3 or less samples
#use Stelligera (use taxa harmonization code)
#facet_wrap by datasetid
```


```{r}
states <- purrr::map(.$siteid, function(y) {
                                      site <- get_sites(y)
                                      dls <- get_downloads(site)
                                      umm@sites[[1]]@collunits[[1]]@colldate
                                      })
            as.data.frame(x,(unlist(states))) %>%
            rename(siteid = x) %>%
            print()

################################################################################

publicationss <- purrr::map(datasetidss, function(y) {
                                      pubs <- get_publications(datasetid = y) %>%
                                        as.data.frame() %>%
                                        .$citation %>%
                                        .[1]
                                       Pubs_short <- substr(pubs,1,100)
})
  
publication_df <- unlist(publicationss) %>%
colnames() <- "publication"

uhh <- as.data.frame(datasetidss,(unlist(publicationss)))
 colnames(uhh) <- c("publication", "datasetid")

                                     
datasetidss <- head(unique(some_diatoms$datasetid),7)   

```

```{r}
library(plyr)
library(gridExtra)
library(ggforce)

 ggplot_stelligera <- okayy_fin_2 %>%
  ggplot(aes(y = round(age, digits = 0), x = count_pct)) +
  ggtitle("Discostella stelligera") + #adds title
  #geom_path() + #adds lines between each value
  #geom_point() + #adds points
  geom_bar(stat = "identity") + # adds bars
  coord_flip() + # makes bars horizontal
  geom_hline(yintercept=100) +
  facet_wrap_paginate(~sitename + datasetid, #function that splits facets into multiple pages
                      nrow = 2, 
                      ncol = 8, 
                      scales = "free_x") + #`facet_wrap()` is what each graph will represent. `scales` is used to give each graph a its own scaled axis (default will scale every graph to same x axis). "free_x" and "free_y" are used seperately to seperate the x axis and y axis respectively.
  scale_y_reverse() +
  labs(x = "Abundance (%)", y =
         #"Age (cal yr BP)"
         "Age") +
  theme_bw()

#RUN ABOVE CODE BEFORE DOING BELOW



my_list <- list() #empty list needed to put loop output to a PDF

required_n_pages <- n_pages(ggplot_stelligera)

for(i in 1:required_n_pages){
  ggplot_stelligera <- okayy_fin_2 %>%
  ggplot(aes(y = age, x = count_pct)) +
  ggtitle("Discostella stelligera") + #adds title
  #geom_path() + #adds lines between each value
  #geom_point() + #adds points
  geom_bar(stat = "identity") + # adds bars
  coord_flip() + # makes bars horizontal
  geom_hline(yintercept=100) + #adds horizontal line
  facet_wrap_paginate(~sitename + datasetid, #function that splits facets into multiple pages
                      nrow = 2, 
                      ncol = 8, 
                      page = i, 
                      scales = "free_x") + #`facet_wrap()` is what each graph will represent. `scales` is used to give each graph a its own scaled axis (default will scale every graph to same x axis). "free_x" and "free_y" are used seperately to seperate the x axis and y axis respectively.
  scale_y_reverse() +
  labs(x = "Abundance (%)", y =
         #"Age (cal yr BP)"
         "Age") +
  theme_bw()

  my_list[[i]] <- ggplot_stelligera #puts each output into a list for PDF function.
  print(ggplot_stelligera) 
}
  
ggsave(
   filename = "plots12.pdf", 
   plot = marrangeGrob(my_list, nrow=1, ncol=1), 
   width = 9, height = 15
)

```

```{r}
#https://stackoverflow.com/questions/8751497/latitude-longitude-coordinates-to-state-code-in-r
latlong_df <- data.frame(x = c(-102.5244,-100.62), y = c(41.77056, 42.626))


library(sp)
library(maps)
library(maptools)

lonlat_to_state_sp <- function(pointsDF) {
    # Prepare SpatialPolygons object with one SpatialPolygon
    # per state (plus DC, minus HI & AK)
    states <- map('county', fill=TRUE, col="transparent", plot=FALSE) #change county to state to get only states
    IDs <- sapply(strsplit(states$names, ":"), function(x) x[1])
    states_sp <- map2SpatialPolygons(states, IDs=IDs,
                     proj4string=CRS("+proj=longlat +datum=WGS84"))

    # Convert pointsDF to a SpatialPoints object 
    pointsSP <- SpatialPoints(pointsDF, 
                    proj4string=CRS("+proj=longlat +datum=WGS84"))

    # Use 'over' to get _indices_ of the Polygons object containing each point 
        indices <- over(pointsSP, states_sp)

    # Return the state names of the Polygons object containing each point
    stateNames <- sapply(states_sp@polygons, function(x) x@ID)
    stateNames[indices]
}

lonlat_to_state_sp(latlong_df)

###############################################################################
library(geonames)
readLines(url("http://api.geonames.org/",open="r"))
options(geonamesUsername= "eomalley") #Note you have to create a username one the website AND enable webservices on your geonames user account at https://www.geonames.org/manageaccount.

options(geonamesUsername="eomalley")

latlong_alt <- GNsrtm3(latlong_df$y[13],latlong_df$x[13])

latlong_alt
```

# 1. ggplot (scatterplot) -> y axis is % abun, x axis is WC characteristic (each facet is a different WC variable)
 -- try using all diatom cores in NE US as dataset.
```{r}
# polygon obtained from http://geojson.io/#map=5/42.049/-75.894
ne <- '{"type": "Polygon",
        "coordinates": [[
            [-80.5517578125,
              39.70718665682654],
            [-66.7529296875,
              39.70718665682654],
            [-66.7529296875,
              47.517200697839414],
            [-80.5517578125,
              47.517200697839414],
            [-80.5517578125,
              39.70718665682654]]]}'

diat_samp <- neotoma2::get_datasets(loc = ne, datasettype = "diatom surface sample",limit = 300) %>%
  get_downloads(., all_data = TRUE) %>%
  samples() %>%
  group_by(sampleid) %>%
  mutate(count_pct = value*100 / sum(value)) %>%
  select(units,value,count_pct,variablename,sampleid,datasetid,siteid,sitename,datasettype) #removing columns to make looking at the data easier, not needed

new_diat_samp_2 <- diat_samp %>%
  mutate(variablename = replace(variablename, 
                                stringr::str_detect(variablename, "stelligera.*"), 
                                "Discostella stelligera"))

new_diat_samp_3 <- new_diat_samp_2 %>%
  dplyr::filter(variablename == "Discostella stelligera") %>%
  
new_diat_samp_4 <- new_diat_samp_3 %>%
  group_by(siteid) %>%
  mutate(., avg_abun = mean(count_pct)) %>%
  distinct(avg_abun, .keep_all = TRUE) %>%
  select(-value,-count_pct)

diat_sites <- unique(new_diat_samp_4$siteid)

wc_samp_data <- get_sites(diat_sites) %>%
  get_datasets() %>%
  neotoma2::filter(datasettype == "water chemistry")

wc_samples <- get_downloads(wc_samp_data) %>%
  samples() %>%
  select(units,value,variablename,sampleid,datasetid,siteid,sitename,datasettype)

wc_samples$avg_abun = NA
new_diat_samp_4$value = NA

new_diat_samp_5 <- new_diat_samp_4 %>%
  select(siteid,variablename,avg_abun)

colnames(new_diat_samp_5) <- c("siteid", "taxa", "avg_abun")

all_data_alt <- merge(wc_samples,new_diat_samp_5)

all_data <- rbind(wc_samples,new_diat_samp_4) %>%
  arrange(.,siteid)
###############################################################################

ggplot_wc_stel <- all_data_alt %>%
  ggplot(aes(y = avg_abun, x = value)) +
  ggtitle("Discostella stelligera") + #adds title
  geom_path() + #adds lines between each value
  geom_point() + #adds points
  facet_wrap_paginate(~variablename, scales = "free_x") + #`facet_wrap()` is what each graph will represent. `scales` is used to give each graph a its own scaled axis (default will scale every graph to same x axis). "free_x" and "free_y" are used seperately to seperate the x axis and y axis respectively.
  labs(x = "WC", y =
         #"Age (cal yr BP)"
         "Abun") +
  theme_bw()


my_list <- list() #empty list needed to put loop output to a PDF

required_n_pages <- n_pages(ggplot_wc_stel)

for(i in 1:required_n_pages){
  ggplot_wc_stel <- all_data_alt %>%
  ggplot(aes(y = avg_abun, x = value)) +
  ggtitle("Discostella stelligera") + #adds title
  #geom_path() + #adds lines between each value
  geom_point() + #adds points
  geom_smooth(method=lm, se=FALSE, col='red', size=1) +
  #geom_bar(stat = "identity") + # adds bars
  #coord_flip() + # makes bars horizontal
  #geom_hline(yintercept=100) + #adds horizontal line
  facet_wrap_paginate(~variablename, #function that splits facets into multiple pages
                      nrow = 4, 
                      ncol = 4, 
                      page = 2, 
                      scales = "free") + #`facet_wrap()` is what each graph will represent. `scales` is used to give each graph a its own scaled axis (default will scale every graph to same x axis). "free_x" and "free_y" are used seperately to seperate the x axis and y axis respectively.
  #scale_y_reverse() +
  labs(x = "WC", y =
         #"Age (cal yr BP)"
         "Abun") +
  theme_bw()

  my_list[[i]] <- ggplot_stelligera #puts each output into a list for PDF function.
  print(ggplot_stelligera) 
}
```

```{r}
#[source](https://open.neotomadb.org/EPD_binder/simple_workflow.html)
#[worldclim format](https://worldclim.com/formats1)
library(raster)

#STEP 1: desired count data
ne <- '{"type": "Polygon",
        "coordinates": [[
            [-80.5517578125,
              39.70718665682654],
            [-66.7529296875,
              39.70718665682654],
            [-66.7529296875,
              47.517200697839414],
            [-80.5517578125,
              47.517200697839414],
            [-80.5517578125,
              39.70718665682654]]]}'

diat_samp_1 <- neotoma2::get_datasets(loc = ne, datasettype = "diatom surface sample",limit = 300) %>%
  get_downloads(., all_data = TRUE) %>%
  samples() %>%
  filter(age < 50)

#######################################################RASTER
spatial <- sf::st_as_sf(diat_samp_1, 
                        coords = c("long", "lat"),
                        crs = "+proj=longlat +datum=WGS84")

worldTmean <- raster::getData('worldclim', var = 'tmean', res = 10)
#worldBio <- raster::getData('worldclim', var = 'bio', res = 10)

spatial$tmean7 <- raster::extract(worldTmax, spatial)[,7] #This adds a column to the `data.frame` `spatial`, that contains the mean July temperature for each taxon at each site (all taxa at a site will share the same value).

meansamp <- spatial %>% 
  dplyr::group_by(siteid) %>% 
  dplyr::summarise(tmean7 = mean(tmean7), .groups = 'keep')

#######################################################PLOT

ggplot_NE_taxa_tempdens <- ggplot() +
  geom_density(data = spatial,
               aes(x = round(tmean7 / 10, 0)), col = 2) +
  facet_wrap(~variablename) +
  geom_density(data = meansamp, aes(x = tmean7 / 10)) +
  xlab("Mean July Temperature") +
  ylab("Kernel Density")


```