---
title: "Workshop November"
author: "Eric O'Malley"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, message=FALSE}
library(neotoma2)
library(dplyr)
library(purrr)
library(sf)
library(tidyverse)
library(rioja)
library(leaflet)
library(leaflegend)
library(ggforce)
```


Diatom taxa that I am interested in working on are Cyclotella stelligera + Discotella stelligera ( all %stelligera%), Aulacoseira ambigua + Melosira ambigua, Aulacoseira granulata + Melosira granulata, and Fragilaria crotonensis.

E - Distribution of one or more taxa - map - size of symbol represents abundance

**PROBLEM: error in retrieving `samples()` when there is multiple default chronologies.

```{r data retrieval, warning=FALSE}
# polygon obtained from http://geojson.io/#map=5/42.049/-75.894
ne <- '{"type": "Polygon",
        "coordinates": [[
            [-80.5517578125,
              39.70718665682654],
            [-66.7529296875,
              39.70718665682654],
            [-66.7529296875,
              47.517200697839414],
            [-80.5517578125,
              47.517200697839414],
            [-80.5517578125,
              39.70718665682654]]]}'


ne_datasets <- neotoma2::get_datasets(loc = ne[1], datasettype = "diatom surface sample", all_data = TRUE)

ne_dl <- get_downloads(ne_datasets, all_data = TRUE)

allSamp <- samples(head(ne_dl,320))


ne_dl_new <- ne_dl[-321]#problem sites
ne_dl_new_2 <- ne_dl_new[-622]#problem sites

allSamp <- samples(head(ne_dl_new_2,800))

length(unique(allSamp$siteid))
view(allSamp)

```

```{r percent abundance function, echo=FALSE}
pc_abun <- function(dl_object) {
  
  samp <- samples(dl_object)
  taxa <- samp$variablename
  count <- samp$value
  sitename <- samp$sitename
  datasetid <- samp$datasetid
  siteid <- samp$siteid
  sampleid <- samp$sampleid
  lat <- samp$lat
  long <- samp$long
  
  
  countDF <- data.frame(lat, long, siteid, sitename, datasetid, sampleid, taxa, count) %>%
    group_by(sampleid) %>%
    mutate(count_pct = count*100 / sum(count))
  return(countDF)
}
```

```{r, warning=FALSE}
diatom_taxa <- pc_abun(ne_dl_new_2) %>%
  dplyr::filter(taxa == "Fragilaria crotonensis") %>%
  arrange(siteid)
```

```{r, out.width = '100%'}
# https://roh.engineering/posts/2021/05/map-symbols-and-size-legends-for-leaflet/

numPal <- colorNumeric('viridis', (diatom_taxa$count_pct))
leaflet(diatom_taxa) %>%
  addTiles() %>%
  addSymbolsSize(values = ~count_pct,
                 lat = ~lat, 
                 lng = ~long,
                 shape = 'circle',
                 color = ~numPal(count_pct),
                 fillColor = ~numPal(count_pct),
                 opacity = .5,
                 baseSize = 7,
                 popup = paste0("<b>", diatom_taxa$sitename,
                                              "</b><br><b>% Abundance:</b> ",
                                              diatom_taxa$count_pct,
                                              "<br><a href=http://apps.neotomadb.org/explorer/?siteids=",
                                              diatom_taxa$siteid,
                                              ">Explorer Link</a>"),
                               options = markerOptions(riseOnHover = TRUE)) %>%
  addLegendSize(
    values = ~count_pct,
    pal = numPal,
    title = '% abundance of F. crotonensis',
    baseSize = 7,
    shape = 'circle',
    orientation = 'horizontal',
    opacity = .5,
    fillOpacity = .3,
    position = 'bottomright',
    breaks = 5)
```







E - Distribution of taxon along environmental gradients; multi-facet - water chem chars
```{r, two variable leaflet}
#This is using the "EVERYTHING" sheet in the aggregate dataset as of 7-20-22
DREXEL_agg <- read.csv("Aggregate Datasets - Diatoms - for Neotoma 7-29-22.csv")
colnames(DREXEL_agg) <- c("drexelid","siteid","sitename","diatom_datasetid", "wc_datasetid")

with_wc_agg <- DREXEL_agg[!is.na(DREXEL_agg$wc_datasetid),]

diatom_datasetid <- tail(with_wc_agg$diatom_datasetid,400) #for some reason I got an error trying to do `head()` instead of `tail()`. Possible wrong datasetid towards beginning.

wc_datasetid <- tail(with_wc_agg$wc_datasetid,400)

diatom_dl <- get_datasets(diatom_datasetid,all_data = TRUE) %>%
  get_downloads()

wc_dl <- get_datasets(wc_datasetid,all_data = TRUE) %>%
  get_downloads()

diat_taxa <- pc_abun(diatom_dl) %>%
  dplyr::filter(taxa == "Fragilaria crotonensis") %>%
  arrange(siteid)

diat_sites <- diat_taxa$siteid

wc_new <- get_sites(diat_sites) %>%
  get_datasets(., datasettype = 'water chemistry') %>%
  get_downloads()

wc_variable <- samples(wc_new) %>%
  dplyr::filter(element == "field") %>%
  dplyr::select('siteid', 'variablename','value') %>%
  dplyr::filter(variablename == "pH") %>%
  arrange(siteid)

wc_diatom <- merge(diat_taxa, wc_variable)
```

```{r}

sumn <- get_datasets(datasettype = "diatom surface sample", all_data=TRUE) %>%
  get_downloads() %>%
  pc_abun() %>%
  dplyr::filter(taxa == "Fragilaria crotonensis") %>%
  arrange(siteid)

diat_sites <- unique(diatom_taxa$siteid)

sumn_wc <- get_sites(diat_sites) %>%
  get_datasets() %>%
  neotoma2::filter(datasettype == "water chemistry") %>%
  get_downloads()

wc_variable <- samples(sumn_wc) %>%
  dplyr::select('siteid', 'variablename','value') %>%
  dplyr::filter(variablename == "Total Phosphorus (unfiltered)") %>%
  arrange(siteid)

diatom_taxa %>%
dplyr::filter(depth < 1.1) %>%
view()

wc_diatom <- merge(diatom_taxa, wc_variable)
```




**OLD**
```{r, editing plotleaflet function, out.width = '100%'}
getColor <- function(wc_diatom) {
  sapply(wc_diatom$value, function(value) {
  if(value <= 8.2) {
    "red"
  } else {
    "blue"
  }})
}

symbols <- makeSymbolIcons(values = wc_diatom$count_pct,
                          shape = 'circle',
                          color = 'black',
                          fillColor = getColor(wc_diatom),
                          width = wc_diatom$count_pct,
                          #pal = numPal,
                          opacity = .7,
                          fillOpacity = .5,
                          baseSize = 15)
            
             map1 <- leaflet(wc_diatom) %>%
              addProviderTiles(providers$Stamen.TerrainBackground) %>%
              addTiles() %>%
              addMarkers(icon = symbols, lng = wc_diatom$long, lat = wc_diatom$lat,
                               popup = paste0("<b>", wc_diatom$sitename,
                                              "</b><br><b>pH:</b> ",
                                              wc_diatom$value,
                                              "<br><a href=http://apps.neotomadb.org/explorer/?siteids=",
                                              wc_diatom$siteid,
                                              ">Explorer Link</a>"),
                               options = markerOptions(riseOnHover = TRUE)) %>%
              addLegendSize(values = wc_diatom$count_pct,
                fillColor = c('white'),
                color = 'black',
                #pal = numPal,
                title = "A. minutissima % Abundance",
                baseSize = 15,
                shape = 'circle',
                orientation = 'horizontal',
                opacity = .7,
                fillOpacity = .3,
                position = 'bottomright',
                breaks = 5) #%>%
               # addLegend(values = c())
             
map1
```



**NEW**
```{r}
getColor <- function(wc_diatom) {
  sapply(wc_diatom$value, function(value) {
  if(value <= 8.2) {
    "red"
  } else {
    "blue"
  }})
}
leaflet(wc_diatom) %>%
  addTiles() %>%
  addSymbolsSize(values = ~count_pct,
                 lat = ~lat, 
                 lng = ~long,
                 shape = 'plus',
                 color = 'black',
                 fillColor = ~getColor(value),
                 opacity = .5,
                 baseSize = 10) %>%
  addLegendSize(
    values = wc_diatom$count_pct,
    color = 'black',
    title = 'Depth',
    shape = 'plus',
    orientation = 'horizontal',
    opacity = .5,
    fillOpacity = 0,
    breaks = 5,
    position = 'bottomright') %>%
  addLegendNumeric(
    pal = numPal, 
    title = 'Magnitude',
    shape = 'stadium',
    values = quakes$mag, 
    fillOpacity = .5,
    decreasing = TRUE,
    position = 'bottomright')

```

```{r}
numPal <- colorNumeric('magma', wc_diatom$value)
leaflet(wc_diatom) %>%
  addTiles() %>%
  addSymbolsSize(values = ~count_pct,
                 lat = ~lat, 
                 lng = ~long,
                 shape = 'circle',
                 color = 'black',
                 fillColor = ~numPal(value),
                 opacity = .5,
                 baseSize = 10) %>%
  addLegendSize(
    values = wc_diatom$count_pct,
    color = 'black',
    title = '% Abundance F. crotonensis',
    shape = 'circle',
    orientation = 'horizontal',
    opacity = .5,
    fillOpacity = 0,
    breaks = 5,
    position = 'bottomright') %>%
  addLegendNumeric(
    pal = numPal, 
    title = 'Total Phosphorus',
    shape = 'stadium',
    values = wc_diatom$value, 
    fillOpacity = .5,
    decreasing = TRUE,
    position = 'bottomright')

```



E - Distribution of taxa along climate gradients; multi-facet - multiple taxa vs temp




E - Strat graphs of taxa in cores - Many cores, one or a few taxa